---
title: "TOD_Multimodal_analysis"
author: "Bosi Chen"
date: "2/22/2022"
output:   
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
library(float)
library(tidyverse)
library(knitr)
library(kableExtra)
library(plyr)
library(dplyr)
library(ggplot2)
library(ppcor)
library(psych)
library(tableone)
library(pander)
library(PMA)
library(tidyr)

options(digits = 4, scipen = 999)

knitr::opts_chunk$set(warning=FALSE, message=FALSE, error = FALSE)
```

```{r, data-readin, include=FALSE, echo=FALSE, warning=FALSE}
# Read in dataset
rm(list=ls()) ## clear the workspace
#dev.off() ## Clear the plots

mypath <- file.path("/Users/bosi/OneDrive_SDSU/JDP/projects/toddler/TOD_T1_ASD")
TOD_T1 = read.csv("/Users/bosi/OneDrive_SDSU/JDP/projects/toddler/TOD_T1_ASD/TOD_T1_Rdata.csv", na.strings = c("", "NA"), header = TRUE)
TOD_T1$dx <- factor(TOD_T1$dx) #1=ASD, 0=TD
TOD_T1$sex <- factor(TOD_T1$sex) #0=Male, 1=Female
TOD_T1_ASD <- TOD_T1[ which(TOD_T1$dx_name=='ASD'),]
TOD_T1_TD <- TOD_T1[ which(TOD_T1$dx_name=='TD'),]

TOD_T1_CCA = read.csv("/Users/bosi/OneDrive_SDSU/JDP/projects/toddler/TOD_T1_ASD/analysis/TOD_T1_aparc_CCA.csv", na.strings = c("", "NA"), header = TRUE)
Brain_struc <- TOD_T1_CCA[,c(9:76)]
Brain_func <- TOD_T1_CCA[,c(77:144)]
SA <- TOD_T1_CCA[,c(9:42)]
fALFF <- TOD_T1_CCA[,c(77:110)]
SA_fALFF <- cbind(SA, fALFF)
SA_fALFF_no2086 <- SA_fALFF[-c(14),]
pheno <- TOD_T1_CCA[,c(147:154)]
pheno_no2086 <- pheno[-c(14),]
```

#Histogram of age distribution 
```{r, Histograms Age, include=TRUE, echo=FALSE, warning=FALSE, fig.width=5, fig.asp=1}

qplot(TOD_T1_ASD$age,
      geom="histogram",
      binwidth =  4,  
      main = "Histogram for Age: ASD", 
      xlab = "Age (Months)",  
      fill=I("red"), 
      col=I("black"),
      ylim = c(0,8),
      xlim=c(10,75))


qplot(TOD_T1_TD$age,
      geom="histogram",
      binwidth =  4,  
      main = "Histogram for Age: TD", 
      xlab = "Age (Months)",  
      fill=I("blue"), 
      col=I("black"),
      ylim = c(0,8),
      xlim=c(10,75))
```

#CCA model of SA/CT with fALFF/ReHo 
```{r, CCA of raw SA/CT with fALFF/RehO, echo = FALSE, warning=FALSE, include=FALSE}
CCA_model <- CCA(Brain_struc, Brain_func, typex = "standard", typez = "standard", K = 1, niter = 15, trace = TRUE, standardize = TRUE)
print(CCA_model)
CCA_permute_model <- CCA.permute(Brain_struc, Brain_func, typex = "standard", typez = "standard", niter = 15, trace = TRUE, nperms = 25, standardize = TRUE)
Brain_struc_CV <- scale(Brain_struc) %*% CCA_model$u
Brain_func_CV <- scale(Brain_func) %*% CCA_model$v
```

##T-test of struc and func CV between groups
```{r, t-tests struc and func CV, echo = TRUE, warning=FALSE, include=TRUE}
t.test(Brain_struc_CV[1:38,], Brain_struc_CV[39:69,])
t.test(Brain_func_CV[1:38,], Brain_func_CV[39:69,])
Brain_struc_CV_diff <- aov(Brain_struc_CV ~dx + age + sex + cnr, data=TOD_T1)
summary(Brain_struc_CV_diff)
Brain_func_CV_diff <- aov(Brain_func_CV ~dx + age + sex + rmsd, data=TOD_T1)
summary(Brain_func_CV_diff)
```

##Linear regression of struc and func CV with age and dx 
```{r, LM of struc and func CV, echo = FALSE, warning=FALSE, include=TRUE}
Struc_age_dx = lm(Brain_struc_CV ~ dx + age + cnr + sex + dx:age, data=TOD_T1)
summary(Struc_age_dx)
ggplot(TOD_T1, mapping = aes(x = age, y = Brain_struc_CV, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

Func_age_dx = lm(Brain_func_CV ~ dx + age + rmsd + sex + dx:age, data=TOD_T1)
summary(Func_age_dx)
ggplot(TOD_T1, mapping = aes(x = age, y = Brain_func_CV, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()
```

##Linear regression of struc and func CV with Mullen ELC 
```{r, LM of struc and funct CV with Mullen ELC, echo = FALSE, warning=FALSE, include=TRUE}
Struc_ELC = lm(Mullen_ELC ~ dx + age + cnr + sex + Brain_struc_CV + Brain_struc_CV:dx, data=TOD_T1)
summary(Struc_ELC)
ggplot(TOD_T1, mapping = aes(x = Brain_struc_CV, y = Mullen_ELC, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

Func_ELC = lm(Mullen_ELC ~ dx + age + rmsd + sex + Brain_func_CV + Brain_func_CV:dx, data=TOD_T1)
summary(Func_ELC)
ggplot(TOD_T1, mapping = aes(x = Brain_func_CV, y = Mullen_ELC, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()
```

##Linear regression of struc and func CV with ADOS CSS
```{r, LM of struc and funct CV with ADOS CSS, echo = FALSE, warning=FALSE, include=TRUE}
Struc_ADOS = lm(ADOS_CSS ~age + cnr + sex + Brain_struc_CV, data=TOD_T1)
summary(Struc_ADOS)
ggplot(TOD_T1, mapping = aes(x = Brain_struc_CV, y = ADOS_CSS)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

Func_ADOS = lm(ADOS_CSS ~age + rmsd + sex + Brain_func_CV, data=TOD_T1)
summary(Func_ADOS)
ggplot(TOD_T1, mapping = aes(x = Brain_func_CV, y = ADOS_CSS)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()
```

#CCA model of SA with fALFF 
```{r, CCA of raw SA with fALFF, echo = FALSE, warning=FALSE, include=FALSE}
CCA_model_SA_fALFF <- CCA(SA, fALFF, typex = "standard", typez = "standard", K = 1, niter = 15, trace = TRUE, standardize = TRUE)
print(CCA_model_SA_fALFF)
CCA_SA_fALFF_weights <- cbind(CCA_model_SA_fALFF[["u"]], CCA_model_SA_fALFF[["v"]])
SA_CV <- scale(SA)%*% CCA_model_SA_fALFF$u
fALFF_CV <- scale(fALFF)%*% CCA_model_SA_fALFF$v
```

##T-test of SA CV and fALFF CV between groups
```{r, t-tests raw SA with fALFF, echo = TRUE, warning=FALSE, include=TRUE}
t.test(SA_CV[1:38,], SA_CV[39:69,])
t.test(fALFF_CV[1:38,], fALFF_CV[39:69,])
SA_CV_diff <- aov(SA_CV ~dx + age + sex + cnr, data=TOD_T1)
summary(SA_CV_diff)
fALFF_CV_diff <- aov(fALFF_CV ~dx + age + sex + rmsd, data=TOD_T1)
summary(fALFF_CV_diff)
```

##Linear regression of SA and fALFF CV with age and dx
```{r, LM of SA and fALFF CV, echo = FALSE, warning=FALSE, include=TRUE}
SA_age_dx = lm(SA_CV ~ dx + age + cnr + sex + dx:age, data=TOD_T1)
summary(SA_age_dx)
ggplot(TOD_T1, mapping = aes(x = age, y = SA_CV, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

fALFF_age_dx = lm(fALFF_CV ~ dx + age + rmsd + sex + dx:age, data=TOD_T1)
summary(fALFF_age_dx)
ggplot(TOD_T1, mapping = aes(x = age, y = fALFF_CV, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()
```

##Linear regression of SA and fALFF CV with Mullen ELC 
```{r, LM of SA and fALFF CV with Mullen ELC, echo = FALSE, warning=FALSE, include=TRUE}
SA_ELC = lm(Mullen_ELC ~ dx + age + cnr + sex + SA_CV + SA_CV:dx, data=TOD_T1)
summary(SA_ELC)
ggplot(TOD_T1, mapping = aes(x = SA_CV, y = Mullen_ELC, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

fALFF_ELC = lm(Mullen_ELC ~ dx + age + rmsd + sex + fALFF_CV + fALFF_CV:dx, data=TOD_T1)
summary(fALFF_ELC)
ggplot(TOD_T1, mapping = aes(x = fALFF_CV, y = Mullen_ELC, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()
```

##Linear regression of SA and fALFF CV with ADOS CSS
```{r, LM of SA and fALFF CV with ADOS CSS, echo = FALSE, warning=FALSE, include=TRUE}
SA_ADOS = lm(ADOS_CSS ~age + cnr + sex + SA_CV, data=TOD_T1)
summary(SA_ADOS)
ggplot(TOD_T1, mapping = aes(x = SA_CV, y = ADOS_CSS)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

fALFF_ADOS = lm(ADOS_CSS ~age + rmsd + sex + fALFF_CV, data=TOD_T1)
summary(fALFF_ADOS)
ggplot(TOD_T1, mapping = aes(x = fALFF_CV, y = ADOS_CSS)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()
```

#Analysis of individual ROIs derived from CCA of SA and fALFF 
##Linear regression of individual ROI SA fALFF with age and dx 
```{r, LM of SA and fALFF ROI, echo = FALSE, warning=FALSE, include=TRUE}
inferiortemporal_area_age_dx = lm(inferiortemporal_area ~ dx + age + cnr + sex + dx:age, data=TOD_T1_CCA)
summary(inferiortemporal_area_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = inferiortemporal_area, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

lateralorbitofrontal_area_age_dx = lm(lateralorbitofrontal_area ~ dx + age + cnr + sex + dx:age, data=TOD_T1_CCA)
summary(lateralorbitofrontal_area_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = lateralorbitofrontal_area, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

parsorbitalis_area_age_dx = lm(parsorbitalis_area ~ dx + age + cnr + sex + dx:age, data=TOD_T1_CCA)
summary(parsorbitalis_area_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = parsorbitalis_area, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

supramarginal_area_age_dx = lm(supramarginal_area ~ dx + age + cnr + sex + dx:age, data=TOD_T1_CCA)
summary(supramarginal_area_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = supramarginal_area, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

caudalmiddlefrontal_fALFF_age_dx = lm(caudalmiddlefrontal_fALFF ~ dx + age + rmsd + sex + dx:age, data=TOD_T1_CCA)
summary(caudalmiddlefrontal_fALFF_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = caudalmiddlefrontal_fALFF, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

fusiform_fALFF_age_dx = lm(fusiform_fALFF ~ dx + age + rmsd + sex + dx:age, data=TOD_T1_CCA)
summary(fusiform_fALFF_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = fusiform_fALFF, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

medialorbitofrontal_fALFF_age_dx = lm(medialorbitofrontal_fALFF ~ dx + age + rmsd + sex + dx:age, data=TOD_T1_CCA)
summary(medialorbitofrontal_fALFF_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = medialorbitofrontal_fALFF, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

rostralmiddlefrontal_fALFF_age_dx = lm(rostralmiddlefrontal_fALFF ~ dx + age + rmsd + sex + dx:age, data=TOD_T1_CCA)
summary(rostralmiddlefrontal_fALFF_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = rostralmiddlefrontal_fALFF, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()

superiorfrontal_fALFF_age_dx = lm(superiorfrontal_fALFF ~ dx + age + rmsd + sex + dx:age, data=TOD_T1_CCA)
summary(superiorfrontal_fALFF_age_dx)
ggplot(TOD_T1_CCA, mapping = aes(x = age, y = superiorfrontal_fALFF, colour = dx_name)) + geom_point() + geom_smooth(method='lm', formula = y~x) + theme_bw()
```
